### 2021-02-08 notebooks/main_analysis.Rmd

---
title: "Main Analyses of the Combined Dataset"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

# Main Analyses of the Combined Dataset

```{r, include=FALSE}
options(warnPartialMatchArgs = FALSE,
        warnPartialMatchDollar = FALSE, 
        warnPartialMatchAttr = FALSE)

knitr::opts_chunk$set(warning = FALSE, message=FALSE)
```

```{r, message=FALSE}

library(nlme)
library(ape)
library(geiger)
library(phytools)
library(broom)
library(xtable)
library(brms)
library(rotl)
library(HDInterval)
library(scales)
library(tidyverse)
library(rstan)
library(loo)
library(viridis)
library(coda)
library(ggrepel)
library(nls.multstart)
source('../R/utilities.r')
source('../R/node_height.r')
set.seed(0)

```

## Overall Plan

This is the main analysis. Publication-quality figures are in
`notebooks/figures` and are each made by their own scripts. This R markdown
notebook uses curated data that's created in `data/` using the
`combined_dataset.r` R script. The sources for this file are all managed by the
`Makefile` in `../data/`.

## Load the Combined Final Merged Dataset

The combined dataset includes data from Leffler et al. (2012), Corbett-Detig
(2015), Stapley et al. (2017), and Romiguier et al. (2014) data. It's all
merged and processed in `data/combined_dataset.r`, which also uses the `taxadb`
and `ritis` to get taxonomical groupings, and the `datelife` package to get a
tree from the Open Tree of Life and calibrate it using datelife.org.

Note that I filter out the parasite/pathogen species because range size
estimates are unreliable (since they depend on host and occurrence data is more
shoddy).

```{r}
d <- read_tsv('../data/combined_data.tsv')

```

Also, I load the Open Tree of Life phylogeny (`otl_tr`) and calibrated datelife
phylogeny (`tr`) created from the `combined_dataset.r` pipeline.

```{r}
load('../data/calibrated_phylogeny.Rdata')
stopifnot(exists('da_tr') && exists('otl_tr') && exists('tr')) 
```

## Curate the main datasets

Our main dataset has a great deal of missing data, since not all taxa have
population sizes, diversity, and map length data. Here, create the main
datasets of the subsequent analyses.

Main datasets (all of metazoan taxa only):

 - `d`: full dataset, with taxa not in phylogeny and missing variables.
 - `dtr`: the subset of the full dataset with taxa in the phylogeny.
 - `da_cc`: all complete cases: map length, diversity, and pop size.
 - `dt_cc`: all complete cases with phylo taxa: map length, diversity, and pop size.
 - `dt_dps`: diversity/population size complete cases with phylo taxa.
 - `da_dps`: diversity/population size complete cases all taxa.

Note: `dt` prefixes are for tree taxa only (`dtr` is used for no suffix, since
`dt` is a function), `da` prefix are for all taxa.

```{r}

# easier to type phyla taxa dataset
dtr <- da_tr
stopifnot(length(tr$tip.label) == nrow(dtr))

# complete data 
da_cc <- d %>% filter(is.finite(log10_popsize), 
                      is.finite(map_length), 
                      is.finite(log10_diversity), 
                      is.finite(log10_body_mass), 
                      is.finite(log10_range))

dt_cc <- dtr %>% filter(is.finite(log10_popsize), 
                        is.finite(map_length), 
                        is.finite(log10_diversity), 
                        is.finite(log10_body_mass), 
                        is.finite(log10_range))
# now, prune the tree
tr_cc <- keep.tip(tr, dt_cc$species)
stopifnot(length(tr_cc$tip.label) == nrow(dt_cc))

# diversity/popsize data
dt_dps <- dtr %>% filter(is.finite(log10_popsize), 
                         is.finite(log10_body_mass), 
                         is.finite(log10_diversity), 
                         is.finite(log10_range))
# now, prune the tree
tr_dps <- keep.tip(tr, dt_dps$species)
stopifnot(length(tr_dps$tip.label) == nrow(dt_dps))

da_dps <- d %>% filter(is.finite(log10_popsize), 
                       is.finite(log10_body_mass), 
                       is.finite(log10_diversity), 
                       is.finite(log10_range))

# map length / popsize data
dt_ml <- dtr %>% filter(is.finite(log10_popsize), 
                        is.finite(log10_body_mass), 
                        is.finite(log10_map_length))
# now, prune the tree
tr_ml <- keep.tip(tr, dt_ml$species)
stopifnot(length(tr_ml$tip.label) == nrow(dt_ml))

da_ml <- d %>% filter(is.finite(log10_popsize), 
                      is.finite(log10_body_mass), 
                      is.finite(log10_map_length))

save(d, dtr, 
     da_cc, dt_cc, tr_cc, 
     dt_dps, tr_dps, da_dps, 
     dt_ml, tr_ml, da_ml, file='../data/main_datasets.Rdata')

```

We lost a few species in matching taxa to the R Open Tree of Life tree; we
quantify that here:

```{r, message=TRUE}

msg <- "Size of animalia in combined dataset, total: %d
                        after matching tree: %d"
message(sprintf(msg, nrow(d), nrow(dtr)))

msg <- "Size of animalia in complete dataset, total: %d
                        after matching tree: %d"
message(sprintf(msg, nrow(da_cc), nrow(dt_cc)))

msg <- "Size of animalia in popsize/diversity dataset, total: %d
                        after matching tree: %d"
message(sprintf(msg, nrow(da_dps), nrow(dt_dps)))

msg <- "Size of animalia in map length/popsize dataset, total: %d
                        after matching tree: %d"
message(sprintf(msg, nrow(da_ml), nrow(dt_ml)))

```

Finally, we load in the posterior draws of the predicted population sizes.

```{r}

load('../data/post_pred_draws.Rdata')

post_popsizes <- post_popsizes %>% 
                  mutate(log10_mean = map_dbl(log10_popsize, mean),
                         log10_sd = map_dbl(log10_popsize, sd))

```

## Diversity Data Source

Is there good mixing across data sources in the combined diversity date set?

```{r}

ggplot(d, aes(log10_popsize, log10_diversity, color=diversity_data_source)) + geom_point()

```

Yes, looks like so.

## Some Map Length, Body Size, and Chromosome Number EDA

Here, are some quick EDA plots. first, chromosome number and the map length
(mostly from Stapley et al data):

```{r}

d  %>% ggplot(aes(hcn, map_length)) + geom_point() + 
  geom_smooth(method='lm') + xlab('haploid chromosome number') + 
  ylab('map length (Morgans)')

```

Body size and map length:

```{r}

d %>% 
  filter_at(vars(hcn, log10_size), ~ !is.na(.)) %>%
  ggplot(aes(log10_size, hcn)) + geom_point() + 
  geom_smooth(method='lm') + ylab('haploid chromosome number') + 
  xlab('log10(size)')

```

Morgans per chromosome by chromosome number:

```{r}
d %>% ggplot(aes(hcn, map_length/hcn)) + 
  geom_point() + 
  geom_smooth(method='lm') + xlab('haploid chromosome number') + 
  ylab('Morgans per chromosome (map length / haploid chromosome number)')

```

What do number of Morgans / chromosome across the data look like?

```{r}

d %>% mutate(morgans_per_chrom = map_length/hcn) %>% 
  pull(morgans_per_chrom) %>% hist(50, main='Morgans / chromosome')

```

## Population Size Biomass Analysis

As a quality control on population size estimates, we can look at humans (we
know they have a modern population size of ~8 billion):

```{r}
d %>% filter(species == 'Homo sapiens') %>% 
  select(log10_popsize, size, log10_density, log10_range)
```

This is low, but not terribly so.

What proportion are terrestrial?

```{r}

prop.table(table(d$is_terrestrial))

```

Based on the predictions of Mora et al. 2011

```{r}

nspecies <- c(terrestrial= 7.770e6, marine=2.15e5)
prop.table(nspecies)

```

Now we do a simple biomass analysis -- back of the envelop type stuff.

```{r}

# https://bionumbers.hms.harvard.edu/bionumber.aspx?&id=107256 
# 18-23% of human body is carbon; seems reasonable this holds
# for other animals

# only complete data
xd <- d %>% filter(is.finite(log10_body_mass))

# in Gt
prop_carbon <- signif(sqrt(0.18 * 0.23), 1)  # ~ 0.2
# 1Gt = 1e12 kg
xd$carbon_mass <- (prop_carbon*(10^xd$pred_log10_body_mass) / 1e3 * 10^xd$log10_popsize) / 1e12

mass_d <- xd %>% 
  # filter so we don't include NA species in nspecies
  filter(is.finite(carbon_mass), is.finite(log10_popsize)) %>%
  group_by(phylum) %>% 
  summarize(total_carbon_mass=sum(carbon_mass, na.rm=TRUE),
            total_log10_popsize=log10(sum(10^log10_popsize, na.rm=TRUE)),
            nspecies = n()) %>%
  mutate(total = sum(total_carbon_mass), prop_carbon_mass = total_carbon_mass / total)

# chordates are incredibly over-represented in my data set by biomass
baron_d <- tribble(~ phylum, ~ bd_mass,
                   'Arthropoda', 0.2 + 1, # terrestrial + marine
                             # fish + livestock + humans + wild mammals + wild birds
                   'Chordata', 0.7 + 0.1 + 0.06 + 0.007 + 0.002,
                   'Annelida', 0.2, 
                   'Mollusca', 0.2, 
                   'Cnidaria', 0.1,
                   'Nematoda', 0.02) %>% 
           mutate(bd_total_biomass = sum(bd_mass), 
                  bd_prop_mass = bd_mass/bd_total_biomass) %>%
           left_join(mass_d, by='phylum')  %>%
           # not we need to reweight the original proportions as some phyla were
           # dropped
           mutate(prop_carbon_mass = prop_carbon_mass / sum(prop_carbon_mass))

ggplot(baron_d, aes(bd_prop_mass, prop_carbon_mass, color=phylum))  + 
  geom_point(size=5) + geom_abline(slope=1, intercept=0) + 
  xlab('Bar-On et al. 2018 proportion') + 
  ylab('My proportion')

```

Arthropods are under sampled here, and chordates are way over sampled (at least
by biomass).

```{r}

ggplot(baron_d) + 
  geom_point(aes(phylum, total_carbon_mass), size=5)

ggplot(baron_d) + 
  geom_point(aes(phylum, bd_mass), size=5)

baron_d <- baron_d %>% mutate(prop_bd_biomass = total_carbon_mass / bd_mass)

baron_d %>%
  select(phylum, prop_bd_biomass)

# using the Bar-On weights
baron_d <- baron_d %>% 
              mutate(reweight = bd_prop_mass / prop_carbon_mass) %>%
              mutate(reweighted_total_carbon_mass = reweight * total_carbon_mass)

total_sample_carbon_mass <- baron_d %>% pull(reweighted_total_carbon_mass) %>% sum()

# total species here
n <- nrow(xd)
total_animal_species <- sum(nspecies)

print(paste0("fraction of animal species in this data set: ", 
             signif(n/total_animal_species, 2)))

# Bar-On et al. (2018) estimate is 2Gt C
# 1e12 kg = 1Gt
total_animal_cmass <- 2

# if the species here were random
signif(total_animal_cmass * n / total_animal_species, 2)
signif(total_sample_carbon_mass, 2)

0.058 / total_animal_cmass

```

Note: initially it was my plan to see if we could get a sense of what fraction
of the total biomass this dataset contains (which is about 3%). This seems
unreasonably high, even after reweighting as above. The problem is that the
3.4e-5 fraction of animal species is obviously not a random sample based on
size / population, and includes humans, domesticated species, etc. -- this
probably is accentuated by the overabundance of chordates compared arthropods
here.

## Stan Model for Map Length and log10 Population Size

Fit the log-normal model, which seems to be a good fit to the data given the
heteroscedasticity.

```{r}
# no social taxa here; excluded

dt_ml_ns <- dt_ml %>% filter(!social)
tr_ml_ns <- keep.tip(tr_ml, dt_ml_ns$species)

# correlation matrix from the tree
ml_cor_ns <- vcv(tr_ml_ns, cor=TRUE)

# data for the log-normal fit
log10_popsize_seq <- seq(5, 18, length.out=100)
ml_dat <- list(N = nrow(dt_ml_ns), 
              log10_popsize = dt_ml_ns$log10_popsize,
              map_length = dt_ml_ns$map_length, 
              cor = ml_cor_ns,
              x_new = log10_popsize_seq)

PHYLO_MM_LOGNORMAL_CACHE <- '../data/phylo_mm_lognormal_chains.Rdata'

if (!file.exists(PHYLO_MM_LOGNORMAL_CACHE)) {
  ln_fit <- stan('../stan/phylo_mm_lognormal.stan', 
                 data=ml_dat, iter=5000, cores=4)
  save(ln_fit, file=PHYLO_MM_LOGNORMAL_CACHE)
} else {
  load(PHYLO_MM_LOGNORMAL_CACHE)
}

print(ln_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r'))
beta <- extract(ln_fit, pars=c('beta'))

mu_ln_fit <- rstan:::extract(ln_fit, 'mu_rep')$mu_rep

save(ln_fit, log10_popsize_seq, mu_ln_fit,
     file='../data/lognorm_popsize_chains.Rdata')

```

Any serious pathologies? Rhats are low; I'm looking for divergent transitions.

```{r}
pairs(ln_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r'))
```

Doesn't look like it.

What is phylogenetic heritability?

```{r}
lambda_est(ln_fit)

```

What is the credible interval?

```{r}
hdi(beta)

```

What does the posterior predictive distribution look like? Social taxa are
outliers (they have higher levels of recombination) and are shown in blue:

```{r}

mu_rep <- extract(ln_fit, pars='mu_rep')$mu_rep
cols <- c('gray32', 'cornflowerblue')
plot(map_length ~ log10_popsize, dt_ml, pch=19, 
     col=cols[factor(dt_ml$social)])
pred  <- colMeans(mu_rep)
lines(log10_popsize_seq, pred)

```

## Continuous Trait Map

Here I generate a bunch of continuous trait maps, and output data for plots.
First, a convenience function.

```{r}

# this simplifies things, since we're working with dataframes
# ordered by the tip labels, and phytools:::contMap requires
# named vectors
contmap <- function(tree, df, col, ...) {
  stopifnot(all(tree$tip.label == df$species))
  x <- setNames(df[[col]], df$species)
  phytools:::contMap(tree, x, ...)
}

```

```{r}

pscm <- contmap(tr_dps, dt_dps, 'log10_popsize', res=200)
dvcm <- contmap(tr_dps, dt_dps, 'log10_diversity', res=200)

save(pscm, dvcm, tr_dps, dt_dps, file='../data/biconmap_data.Rdata')

```

Now, for the map length / population size dataset.

```{r}

psmlcm <- contmap(tr_ml, dt_ml, 'log10_popsize', res=200)
mlcm <- contmap(tr_ml, dt_ml, 'log10_map_length', res=200)

save(mlcm, psmlcm, tr_ml, dt_ml, file='../data/biconmap_maplen_data.Rdata')

```

## Phylogenetic Mixed-Effects Model of Diversity Population Size Relationship

This is a primary relationship we want to fit, accounting for phylogeny:

```{r}

dps_cor <- vcv(tr_dps, corr=TRUE)
log10_popsize_seq <- seq(4, 16, length.out=100)

dps_dat <- list(N = nrow(dt_dps),
          x = dt_dps$log10_popsize,
          y = dt_dps$log10_diversity,
          x_new=log10_popsize_seq,
          cor=dps_cor)

PHYLO_MM_REGRESSION_CACHE <- '../data/phylo_mm_regression_chains.Rdata'
if (!file.exists(PHYLO_MM_REGRESSION_CACHE)) {
  dps_fit <- stan('../stan/phylo_mm_regression.stan', 
                  seed=0,
                  data=dps_dat, iter=5000, cores=4)
  save(dps_fit, file=PHYLO_MM_REGRESSION_CACHE)
} else {
  load(PHYLO_MM_REGRESSION_CACHE)
}

print(dps_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r'))
pairs(dps_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r', 'lp__'))

phi <- extract(dps_fit, pars='phi')$phi
beta <- extract(dps_fit, pars='beta')$beta
dps_mu_rep <- extract(dps_fit, pars='mu_rep')$mu_rep
plot(dps_dat$x, dps_dat$y, pch=19, xlab='log10 pop size', 
     ylab='log10 diversity')
lines(log10_popsize_seq, colMeans(dps_mu_rep))

dps_ols <- lm(log10_diversity ~ log10_popsize, dt_dps)
summary(dps_ols)

```

The chains mixed well. For quality assurance I compare this to a PGLS, to check
my Stan model is working not too differently than a pre-packaged solution.

```{r}

# compare with pgls
summary(gls(log10_diversity ~ log10_popsize, 
            correlation = corBrownian(phy=tr_dps), 
            data=dt_dps, method = "ML"))

```

This is reasonably close to the phylogenetic mixed-effects model.

Now, let's do a simple bootstrap to get the CIs for the OLS fit.

```{r}

fit_fun <- function(x) {
  lm(log10_diversity ~ log10_popsize, data=x)
}

nboot <- 10000
dps_bs <- tibble(i = 1:nboot) %>%
       mutate(data = map(i, ~ bs_samp(dt_dps))) %>%
       mutate(fit = map(data, fit_fun)) %>%
       mutate(slope = map_dbl(fit, ~ coef(.)[2])) %>%
       mutate(intercept = map_dbl(fit, ~ coef(.)[1])) %>%
       # don't save the permuted data sets
       select(-data)

dps_bs_cis <- rbind(ci(dps_bs$intercept, coef(dps_ols)[1]),
                    ci(dps_bs$slope, coef(dps_ols)[2]))
dps_bs_cis
```

Now, save all these data

```{r}

save(dps_fit, dps_ols, dt_dps, dps_bs_cis,
     file='../data/diversity_popsize_chains.Rdata')

```

How many taxa?

```{r}
dps_dat$N
```

Credible intervals:

```{r}
round(hdi(beta), 2)
round(lambda_est(dps_fit), 2)
```

## Diversity Population Size Relationship by Phylum

Next, let's look at phyla separately.

```{r}

min_taxa <- 10
dt_dps_phyla <- dt_dps %>% group_by(phylum) %>%
                  nest() %>% 
                  mutate(n = map_int(data, nrow)) %>%
                  filter(n > min_taxa)

fit_phylo_mm <- function(df, full_tree,
                         stan_file='../stan/phylo_mm_regression.stan', 
                         iter=5000) {
  tr <- keep.tip(tr, df$species)
  cor <- vcv(tr, corr=TRUE)
  log10_popsize_seq <- seq(4, 16, length.out=100)

  dat <- list(N = nrow(df),
              x = df$log10_popsize,
              y = df$log10_diversity,
              x_new=log10_popsize_seq,
              cor=cor)

  fit <- stan(stan_file,
              data=dat, iter=iter, seed=0, cores=4)
  fit
}

# fit all models: 
PHYLO_MM_REGRESSION_BY_PHYLA_CACHE <- '../data/phylo_mm_regression_by_phyla_chainsdf.Rdata'

if (!file.exists(PHYLO_MM_REGRESSION_BY_PHYLA_CACHE)) {
  dt_dps_phyla <- dt_dps_phyla %>% 
                   mutate(fit = map(data, fit_phylo_mm, full_tree=tr_dps))
  save(dt_dps_phyla, file=PHYLO_MM_REGRESSION_BY_PHYLA_CACHE)
} else {
  load(PHYLO_MM_REGRESSION_BY_PHYLA_CACHE)
}

```

Now, we have to ensure convergence worked. For arthropods:

```{r}

dps_pars <- c('phi', 'beta', 'sigma_p', 'sigma_r', 'lp__')

pairs(dt_dps_phyla$fit[[3]], pars=dps_pars)

print(dt_dps_phyla$fit[[3]], pars=dps_pars)

```

This looks fine -- all Rhats are below 1.01 now.

For molluscs:

```{r}

pairs(dt_dps_phyla$fit[[2]], pars=dps_pars)

print(dt_dps_phyla$fit[[2]], pars=dps_pars)

```

Again, all Rhats are below 1.01, though clearly there is a high
degree of correlation between slope and intercept (low information
in data to differentiate these). We have some divergent
transitions but this seems to not indicate any pathologies in the
geometry of the posterior.

For chordates (hardest for last):

```{r}

pairs(dt_dps_phyla$fit[[1]], pars=dps_pars)
print(dt_dps_phyla$fit[[1]], pars=dps_pars)

```

Now, something is has gone wrong; Rhats are high and we have many
divergent transitions. The pairs plot shows many pathologies in
the geometry of the posterior surface. The histogram of `sigma_r`
shows that there's a posterior mode with low residual variance
that has a very high log posterior. This seems to indicate that
all the variance could be explained by phylogenetic effects,  not
residual tips.

This requires we fit a special model without `sigma_r` as a term.

```{r}

alt_stan_file <- '../stan/phylo_mm_regression_noresid.stan'

PHYLO_MM_REGRESSION_BY_PHYLA_ALT_CACHE <- '../data/phylo_mm_regression_phyla_alt_chains.Rdata'

if (!file.exists(PHYLO_MM_REGRESSION_BY_PHYLA_ALT_CACHE)) {
  chor_fit <- fit_phylo_mm(dt_dps_phyla$data[[1]], tr_dps, stan_file=alt_stan_file, iter=2000)
  save(chor_fit, file=PHYLO_MM_REGRESSION_BY_PHYLA_ALT_CACHE)
} else {
  load(PHYLO_MM_REGRESSION_BY_PHYLA_ALT_CACHE)
}

dps_pars_alt <- c('phi', 'beta', 'sigma_p', 'lp__')
pairs(chor_fit, pars=dps_pars_alt)
print(chor_fit, pars=dps_pars_alt)

```

This greatly helped in the model fitting process. Note that
parameter estimates have not changed:

```{r}
hdi(rstan:::extract(chor_fit, 'beta')$beta)
hdi(rstan:::extract(dt_dps_phyla$fit[[1]], 'beta')$beta)

```

Note that while it's tempting to compare log posteriors between
these two models as some sort of model comparison approach, this is
[not
appropriate](https://discourse.mc-stan.org/t/model-comparison-lp--/1849/3).

Now, we do simple OLS estimates and bootstraps, and save everything:

```{r}

dt_dps_phyla <- dt_dps_phyla %>% 
                 mutate(ols = map(data, fit_fun)) %>%
                 mutate(intercept = map_dbl(ols, ~ coef(.)[1]),
                        slope = map_dbl(ols, ~ coef(.)[2]))

nboot <- 10000

chor_bs <- tibble(i = 1:nboot) %>%
       mutate(data = map(i, ~ bs_samp(dt_dps_phyla$data[[1]]))) %>%
       mutate(fit = map(data, fit_fun)) %>%
       mutate(slope = map_dbl(fit, ~ coef(.)[2])) %>%
       mutate(intercept = map_dbl(fit, ~ coef(.)[1])) %>%
       # don't save the permuted data sets
       select(-data)

arth_bs <- tibble(i = 1:nboot) %>%
       mutate(data = map(i, ~ bs_samp(dt_dps_phyla$data[[2]]))) %>%
       mutate(fit = map(data, fit_fun)) %>%
       mutate(slope = map_dbl(fit, ~ coef(.)[2])) %>%
       mutate(intercept = map_dbl(fit, ~ coef(.)[1])) %>%
       # don't save the permuted data sets
       select(-data)

mols_bs <- tibble(i = 1:nboot) %>%
       mutate(data = map(i, ~ bs_samp(dt_dps_phyla$data[[3]]))) %>%
       mutate(fit = map(data, fit_fun)) %>%
       mutate(slope = map_dbl(fit, ~ coef(.)[2])) %>%
       mutate(intercept = map_dbl(fit, ~ coef(.)[1])) %>%
       # don't save the permuted data sets
       select(-data)

dt_dps_phyla$boot <- list(chor_bs, arth_bs, mols_bs)

get_cis <- function(boot, slope, intercept) {
  cis <- ci(boot$slope, slope, with_est=FALSE)
  cii <- ci(boot$intercept, intercept, with_est=FALSE)
  tibble(slope_lower=cis[1], slope_upper=cis[2],
         intercept_lower = cii[1], intercept_upper=cii[2])
}

dt_dps_phyla <- dt_dps_phyla  %>%
      mutate(ci_bs = pmap(list(boot, slope, intercept), get_cis)) %>% unnest(ci_bs) %>%
      select(-boot)

save(dt_dps_phyla, chor_fit, 
     file='../data/phyla_diversity_popsize_chains.Rdata')

# just the OLS
ols_phyla <- dt_dps_phyla %>% select(-fit)
save(ols_phyla, file='../data/phyla_diversity_popsize_ols.Rdata')

```

## Phylogenetic Mixed-Effects Model of Diversity and Body Size Relationship

Is the relationship between diversity and population size driven by body mass?
We fit that here to take a look:

```{r}

log10_body_mass_seq <- seq(min(dt_dps$log10_body_mass),
                       max(dt_dps$log10_body_mass), length.out=100)

# body mass version
dbm_dat <- list(N = nrow(dt_dps),
          x = dt_dps$log10_body_mass,
          y = dt_dps$log10_diversity,
          x_new=log10_body_mass_seq,
          cor=dps_cor)

PHYLO_MM_REGRESSION_BODY_SIZE_CACHE <- '../data/phylo_mm_regression_body_size_chains.Rdata'

if (!file.exists(PHYLO_MM_REGRESSION_BODY_SIZE_CACHE)) {
  dbm_fit <- stan('../stan/phylo_mm_regression.stan', 
                  seed=0,
                  data=dbm_dat, iter=5000, cores=4)
  save(dbm_fit, file=PHYLO_MM_REGRESSION_BODY_SIZE_CACHE)
} else {
  load(PHYLO_MM_REGRESSION_BODY_SIZE_CACHE)
}

print(dbm_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r'))
pairs(dbm_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r', 'lp__'))

phi <- extract(dbm_fit, pars='phi')$phi
beta <- extract(dbm_fit, pars='beta')$beta
dbm_mu_rep <- extract(dbm_fit, pars='mu_rep')$mu_rep
plot(dbm_dat$x, dbm_dat$y, pch=19, xlab='log10 body mass', 
     ylab='log10 diversity')
lines(log10_body_mass_seq, colMeans(dbm_mu_rep))

dbm_ols <- lm(log10_diversity ~ log10_body_mass, dt_dps)
summary(dbm_ols)

```

The chains mixed well (Rhats < 1.01 and the pair plots look good). What's the
HDI of beta?

```{r}
hdi(beta)
```

For quality assurance, I run the PGLS and compare parameter estimates. They are
very similar (though it should be noted this model doesn't account for residual
variance):

```{r}
# compare with pgls
summary(gls(log10_diversity ~ log10_body_mass, 
            correlation = corBrownian(phy=tr_dps), 
            data=dt_dps, method = "ML"))

```

## Phylogenetic Mixed-Effects Model of Diversity and Range

Next, we look to see if the diversity population size relationship is driven by
another variable used in population size estimates: range.

```{r}

log10_range_seq <- seq(min(dt_dps$log10_range),
                       max(dt_dps$log10_range), length.out=100)

# body mass version
drn_dat <- list(N = nrow(dt_dps),
          x = dt_dps$log10_range,
          y = dt_dps$log10_diversity,
          x_new=log10_range_seq,
          cor=dps_cor)

PHYLO_MM_REGRESSION_RANGE_CACHE <- '../data/phylo_mm_regression_range_chains.Rdata'
if (!file.exists(PHYLO_MM_REGRESSION_RANGE_CACHE)) {
  drn_fit <- stan('../stan/phylo_mm_regression.stan', 
                  seed=0,
                  data=drn_dat, iter=5000, cores=4)
  save(drn_fit, file=PHYLO_MM_REGRESSION_RANGE_CACHE)
} else {
  load(PHYLO_MM_REGRESSION_RANGE_CACHE)
}

print(drn_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r'))
pairs(drn_fit, pars=c('phi', 'beta', 'sigma_p', 'sigma_r', 'lp__'))

phi <- extract(drn_fit, pars='phi')$phi
beta <- extract(drn_fit, pars='beta')$beta
drn_mu_rep <- extract(drn_fit, pars='mu_rep')$mu_rep
plot(drn_dat$x, drn_dat$y, pch=19, xlab='log10 range', 
     ylab='log10 diversity')
lines(log10_range_seq, colMeans(drn_mu_rep))

drn_ols <- lm(log10_diversity ~ log10_range, dt_dps)
summary(drn_ols)

```

Again, we compare to the PGLS fit:

```{r}

# compare with pgls
summary(gls(log10_diversity ~ log10_range, 
            correlation = corBrownian(phy=tr_dps), 
            data=dt_dps, method = "ML"))

```

And again the estimates are close, but the PGLS slope estimates are weakened a
bit, likely by allowing for an additional source of variance (the diagonal
residual variance).

Now, we save all these fits for yet another supplementary figure.

```{r}

save(drn_fit, dbm_fit, drn_ols, dbm_ols, 
     log10_body_mass_seq, log10_range_seq,
     file='../data/diversity_range_bodymass_chains.Rdata')

```

## Node Height Tests

Now, let's look at node height.

```{r}

psnh <- node_height(dt_dps$log10_popsize, multi2di(tr_dps))
rnnh <- node_height(dt_dps$log10_range, multi2di(tr_dps))
bmnh <- node_height(dt_dps$log10_body_mass, multi2di(tr_dps))
dvnh <- node_height(dt_dps$diversity, multi2di(tr_dps))

save(psnh, rnnh, bmnh, dvnh, file='../data/nodeheights.Rdata')

```

Just to ensure there's nothing weird going on, let's simulate
Brownian motion 1000 times and see how many times we see a
significant relationship:

```{r}

mtr <- multi2di(tr_dps)
nhsims <- tibble(i=1:5000) %>% 
           mutate(bm_trait = map(i, ~ fastBM(mtr))) %>%
           mutate(nh_rlm = map(bm_trait, ~ node_height(., mtr))) %>%
           mutate(nh_lm = map(bm_trait, ~ node_height(., mtr, method='lm')))

get_lm_pval <- function(x) {
  summary(x$fit)$coefficients[2, 4]
}

get_rlm_pval <- function(x) {
  tval <- summary(x$fit)$coefficients[2, 3]
  df <- summary(x$fit)$df[2]
  dt(abs(tval), df)
}

nhsims %>% mutate(pval = map_dbl(nh_rlm, get_rlm_pval)) %>%
  ggplot(aes(x=pval)) + geom_histogram()

```

This is the distribution of p-values under robust regression. This shows the null-model is a bit wrong (overly conservative); see [David Robinson's great article on this](http://varianceexplained.org/statistics/interpreting-pvalue-histogram/). This is a bit concerning, but note that the p-value histogram under the linear model looks better:

```{r}

nhsims %>% mutate(pval = map_dbl(nh_lm, get_lm_pval)) %>%
  ggplot(aes(x=pval)) + geom_histogram(binwidth=0.02)

```

So the odd `rlm()` p-value distribution seems to be an artifact of
the robust regression.

Overall, simulating Brownian motion shows that there isn't a
strong reason to suspect the shape of this phylogeny affects the
node-height test.

What does a random BM node height test?

```{r}
plot_nh(nhsims$nh_rlm[[2]])
```

```{r}

# no drosophila
dt_dps_nd <- dt_dps %>% filter(genus != 'Drosophila')
tr_dps_nd <- keep.tip(tr_dps, dt_dps_nd$species)

# diversity
dvnh_nd <- node_height(dt_dps_nd$diversity,
                       multi2di(tr_dps_nd))
plot_nh(dvnh_nd)
summary(dvnh_nd$fit)

# popsize
psnh_nd <- node_height(dt_dps_nd$log10_popsize, 
                    multi2di(tr_dps_nd))
plot_nh(psnh_nd)

# 
outliers <- dvnh_nd$d  %>% filter(pic > 0.001) %>% pull(tips) 
not_outliers <- dvnh_nd$d  %>% filter(pic < 0.001) %>% pull(tips)

outliers_ranges <- sapply(outliers, function(x) diff(range(dt_dps_nd$log10_range[dt_dps_nd$species %in% x])))

not_outliers_ranges <- sapply(not_outliers, function(x) diff(range(dt_dps_nd$log10_range[dt_dps_nd$species %in% x])))

tibble(range = c(not_outliers_ranges, outliers_ranges), 
       group = c(rep('outlier', length(not_outliers_ranges)), 
                 rep('not outlier', length(outliers_ranges)))) %>%
  ggplot(aes(x=group, y=range)) + 
geom_boxplot()

  geom_histogram(position='dodge')

```

## Genome Size

A quick EDA plot of log genome size and log population size:

```{r}

d %>% filter(species != "Ambystoma tigrinum", 
             is.finite(log10_popsize),
             is.finite(genome_size)) %>%
  ggplot(aes(log10_popsize, genome_size)) + geom_point() + 
    scale_y_log10()

```

A simple PGLS of log genome size and log population size:

```{r}

# TODO

```

## Simple Hitchhiking Effects Model based on Elyashiv et al. (2016)

Based on the model of Coop and Ralph (2012) and Elyashiv et al. (2016).

Elyashiv et al. 2016's selection coefficient distribution, based on
Supplementary Table S1.

```{r}

# Table S1, with "missing substitutions"; introns/intergenic are all NA
# so not included (same as Table S11)
dm_s <- tribble(~ log10_s, ~ feature, ~ alpha,
                     -1.5,   'exon',       NA,
                     -1.5,   'utr',        NA, 
                     -2.5,   'exon',       0.6,
                     -2.5,   'utr',        NA,
                     -3.5,   'exon',       3.5,
                     -3.5,   'utr',        NA,
                     -4.5,   'exon',       NA,
                     -4.5,   'utr',        5.1,
                     -5.5,   'exon',       36.3,
                     -5.5,   'utr',        42.1) %>% 
         mutate(alpha = alpha/100., s=10^log10_s) # convert from percentage

```

Next are the substitution rates, from Supplementary Text Section A.

```{r}

# Substitution rates
# These are taken from Supplementary Text Section A, p. 5
# Original source: Hu et al. 
# exon is nonsyn
dm_subs <- tribble(~ feature,       ~ prop_data, ~ num_subs,  ~ rate,
                         'exon',            0.7,      64205,  0.0092,
                          'utr',           0.45,     153765,   0.025,
                       'intron',           0.43,     456401,   0.032)

dm_subs <- tribble(~ feature,    ~ num_subs,  ~ rate,
                      'exon',         64205,  0.0092,
                       'utr',        153765,   0.025,
                    'intron',        456401,   0.032)

```

Next, to get a sense of the proportion of base pairs in each annotation
feature, I load in the summary created  in `../data/dmel/`. Note that these are
only for the chromosomes in the Elyashiv et al. study.

```{r}
# annotation 
dm_ann <- read_tsv('../data/dmel/dmel_annotation_summary.tsv')
dm_sl <- read_tsv('../data/dmel/dmel_seqlens.tsv')

# total genome length 
G <- sum(dm_sl$length)

# genic total bp, including CDS, introns, UTRs
Ggenic <- dm_ann %>% filter(feature %in% c('cds', 'intron', 'UTR')) %>% 
  pull(total_bp) %>% sum()

```

For my approximations, I use Drosophila melanogaster values of $\alpha$ and
$m/T$. The main thing is $J_{2,2}$ based on (1) exogenous $N$ and $L$ and (2)
the inferred Elyashiv et al. Drosophila distribution of selection coefficients.

I first average over the features to get a marginalized selection coefficient
distribution. I weight by proportions of UTRs to exonic sequences in the
annotation.

```{r}

# dm_seldist_simple <- tribble(~ log10_s, ~ prop,
#                                   -3.5,    0.04,
#                                   -5.5,     0.4)

dm_seldist<- dm_s %>% left_join(dm_ann)  %>%
  group_by(feature) %>%
  # Total alpha by feature. Sanity check: these match Table S6
  mutate(alpha_feature = sum(alpha, na.rm=TRUE)) %>%
  # calc the selection coefficient distribution by feature 
  mutate(g_feature = alpha / alpha_feature)  %>%
  filter(is.finite(g_feature)) %>%
  ungroup() %>%
  group_by(log10_s, s) %>%
  summarize(g = weighted.mean(g_feature, prop_genome, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(g = g/sum(g))

dm_seldist

```

Now, I estimate the $\alpha$, averaging across annotation features.

```{r}

# alpha is the fraction of substitutions (in genic regions) that are beneficial
dm_alphas <- dm_s %>% left_join(dm_ann) %>% 
               group_by(feature, prop_genome) %>% 
               summarize(alpha = sum(alpha, na.rm=TRUE), 
                         total_bp = unique(total_bp)) %>%
               ungroup() %>%
               mutate(alpha_mean = weighted.mean(alpha, prop_genome))

dm_alpha <- unique(dm_alphas$alpha_mean)
stopifnot(length(dm_alpha) == 1)

dm_alphas

```

These values seem entirely reasonable given Table 1 of Sella et al. (2009).

Next I estimate $m$, the number of substitutions occurring in coding regions.
This uses Elyashiv et al.'s Supplementary Text which reports the substitutions
per feature (input above as `dm_subs`). Again, this is just for the automsomal
chromosomes.

```{r}

# total_bps is the number of bps per feature 
dm_ms <- dm_subs %>% left_join(dm_ann) %>%
          # note, we exclude introns since Elyashiv found 
          # no evidence of adaptive substitutions in introns, 
          # but see note on p. 13 2nd to last paragraph.
          filter(feature %in% c('exon', 'utr')) %>%
          mutate(m = rate * total_bp)

dm_m <- sum(dm_ms$m)

dm_ms %>% select(feature, num_subs, rate, total_bp, m, prop_genome)
dm_m

```

Next, we sanity check this. To get an idea of whether this is reasonable, we
need the divergence times between *D. melanogaster* and *D. simulans*.

```{r}
# parameters
dmel_sim_div <- 4.2e6 # Mya, from Obbard

# assuming Ne of D. mel is 1e6
dm_mu <- 2.8e-9
dm_Ne <- 1e6

dm_pi = da_cc %>% filter(species == "Drosophila melanogaster") %>% pull(diversity)
# 0.2 from syn subs per site in Supp. p. 5
dmel_sim_div2 <- (0.2 / 2) / dm_pi * dm_Ne # Mya, from method in Elyashiv

# I use Obbard's -- both are similar
T <- dmel_sim_div * 10 # gens

dm_m / (2*T * sum(dm_ms$total_bp))

```

This substitution rate is a bit lower than the mutation rate, but given these
are genic regions, this is not too surprising. Overall, nothing too out of the
ordinary here.

Finally, we work to get an estimate of $J_{2,2}$. Here I explore two
approaches, one a simplified version of Elyashiv et al. where I use their
inferred selection coefficient distribution, and another with fixed $s$. First,
some functions:

```{r}

# fixation time, based on diffusion theory (Elyashiv et al. eqn 4)
tau <- function(s, N) {
  gamma <- -digamma(1)
  2*(log(4*N*s) + gamma - 1/(4*N*s)) / s
}

# simple J22 with fixed s, without haldane
J22 <- function(s, N, L){
  t <- tau(s, N)
  # from Mathematica 
  (1 - exp(-2*L*t))/(2*t)
}

# J22 for distribution of s
J22s <- Vectorize(function(s, w, N, L) {
  ts <- tau(s, N)
  # Inf out any negative fixation times (when Ns is small)
  if (any(ts < 0)) warning("some fixation times < 0")
  ts[ts < 0] <- Inf
  sum(w * (1 - exp(-2*L*ts))/(2*ts))
}, c('N', 'L'))

```

What is the $\nu_{BP}$? I also extract some other relevant parameters.

```{r}
# implied nu_bp
dm_nu_bp <- dm_alpha * dm_m / (G*2*T)
dm_nu_bp

dm_log10_Nc <- da_ml %>% filter(species == "Drosophila melanogaster") %>% pull(log10_popsize) 
dm_Nc <- 10^dm_log10_Nc

# some other Drosophila params
dm_L <- da_ml %>% filter(species == "Drosophila melanogaster") %>% pull(map_length)
dm_U <- 1.6 # from Elyashiv
dm_rbp <- mean(c(2.14, 2.77, 2.2, 1.97)) * 1e-2 / 1e6 # from Petrov's website

```

What about $J_{2,2}$? Here, I find that my simple model that marginalizes over
the distribution of selection coefficients seems to estimate a $J_{2,2}$ that
is much weaker (two orders of magnitude) than estimates that have been
previously published and compared to the rate of coalescent events
(Supplementary Table S6). Here is the $J_{2,2}$ from the distribution of
selection coefficients:

```{r}
# J22 
# NOTE: taking L = Inf does not matter since ts is so high.
# What drives this is the 1/fixation_time term.
dm_J22_Elyashiv <- J22s(dm_seldist$s,  dm_seldist$g, dm_Ne, Inf)  
dm_J22_Elyashiv

# 2 N nu_bp J22 / rbp
twoSN_Elyashiv <- 2 * dm_Ne * dm_nu_bp / dm_rbp * dm_J22_Elyashiv
twoSN_Elyashiv

2 * dm_nu_bp / dm_rbp * dm_J22_Elyashiv

```

How does this compare to the value published in the Supplementary Table (S6) of
Elyashiv et al?

```{r}

rs <- 0.92 # from Table S6
# implied by the rs = 0.92 value in Supplementary Table S3
dm_J22_Elyashiv2 <- rs / (2 * dm_Ne * dm_nu_bp / dm_rbp)

# implied J22
signif(rs / (2 * dm_Ne * dm_nu_bp / dm_rbp), 3)

```

Let's compare these to the numbers from Sella et al. 2009.

```{r}

# from Sella et al. 2009
sweep_ests <-
  tribble(            ~ study,                  ~ s,          ~ nu,   ~ nu_gamma,
     'Wiehe and Stephan 1992',             NA_real_,      NA_real_,   1.3e-8,
        'Li and Stephan 2006',      (0.2 + 0.5)/200,  (6+9)/2*1e-11,      NA,
            'Andolfatto 2007',                 1e-5,        7.5e-10,    3e-8,
      'Macpherson et al 2007',                 0.01,        3.6e-12,    1e-7,
          'Jensen et al 2008',              0.2/100,          4e-11,    4e-7) %>%
        mutate(twoSN = nu_gamma * 0.075 / dm_rbp) %>%
        mutate(approx_J22 = twoSN / (2 * dm_Ne * nu / dm_rbp))  %>%
        add_row(study="***", s=NA, nu=dm_nu_bp, approx_J22 = dm_J22_Elyashiv2)
sweep_ests

write_tsv(sweep_ests, "../data/sweep_param_ests.tsv")

```

Below we show all of these, mine is `***`:

```{r}

sweep_ests %>%
  mutate(color=ifelse(study == '***', 'red', 'black')) %>%  
  ggplot() + 
  geom_hline(yintercept=0, color='gray32') + 
  geom_point(aes(nu, 0, color=color), size=4) + 
  geom_text_repel(aes(nu, 0, label=study), point.padding=1, box.padding=3, seed=1) +
  scale_x_log10() +   scale_colour_identity()+
  xlab("beneficial substitutions per bp (ν_bp)") + ylab('')

```

Is this reasonable?  This seems roughly in line with previous findings.

How does our $J_{2,2}$ compare to other published estimates of $J_{2,2}$?

```{r}

# dataset from Graham's notes
dshap_raw <- read_tsv('../data/shapiro_et_al_2007.tsv', col_names=FALSE)

# unsure what all the columns are, but this is what matters
dshap <- tibble(rec = dshap_raw$X3, pi = dshap_raw$X4)

hh_nls <- nls(pi ~ neutral.pi*rec/(rec + alpha), dshap, 
              start=list(neutral.pi=2, alpha=1e-9))

plot(pi ~ rec, dshap, pch=19, xlab='recombination (cM/Mb)', 
     ylab=latex2exp::TeX('$\\pi$'))
rec <- seq(0, 2.5, length.out=100)
y <- predict(hh_nls, data.frame(rec=rec))
lines(rec, y, col='red', lwd=3)

pi0 <- coef(hh_nls)['neutral.pi']
alpha <- coef(hh_nls)['alpha']
lines(rec, pi0/(1 + alpha/rec), col='red', lwd=3)
lines(rec, pi0/(1 + twoSN_Elyashiv/rec), col='blue', lwd=3)

lines(rec, pi0/(1 + rs/rec), col='green', lwd=3, lty=2)

legend(0, 0.06, c('non-linear least squares', 
                  latex2exp::TeX("sel coef $J_{2,2}$"),
                  latex2exp::TeX('Elyashiv et al. $r_S$')),
        fill=c('red', 'blue', 'green'),
        bty='n', border=0, cex=1, ncol=1)

# the implied 2NS / rbp is below -- this is the same 
# as in Coop and Ralph (2012) Table 1
alpha / (100 * 1e6) # units are changed to cM/Mb

# the J22 is
alpha / (100 * 1e6 * 2*dm_Ne * dm_nu_bp)
dm_J22_Elyashiv

```

Overall the selection coefficient distribution approach underestimates
$J_{2,2}$. Looking at Elyashiv et al., this is because the homogeneous sweep
model spaces substitutions too evenly. I use the empirical $J_{2,2}$ which is
stronger and more realistic. Let's look at some comparisons:

```{r}

sweep_ests %>%
   # implied by cs parameter estimate in supp
   add_row(study="Elyashiv et al. 2012", s=NA, nu=dm_nu_bp, twoSN=0.92, 
           approx_J22=0.92 / (2*dm_Ne * dm_nu_bp / dm_rbp)) %>%
   # roughly the homogeneous sweep model
   mutate(one_over_fix= 0.5 / tau(s, dm_Ne), off=one_over_fix/approx_J22)

```

How do the diffusion and deterministic fixation times compare? Could this be an
issue?

```{r}

# how does our diffusion fixation time compare to the deterministic?
crossing(log10_s = -(1:5), log10_N = 3:10) %>%
  mutate(s=10^log10_s, N = 10^log10_N) %>%
  mutate(`simple` = 4/s * log(2 * N), diffusion=tau(s, N)) %>% 
  filter(diffusion > 0) %>%
  ggplot(aes(simple, diffusion, color=as.factor(log10_s), shape=as.factor(log10_N))) + 
  geom_point() + scale_x_log10() + scale_y_log10() + geom_abline(slope=1, intercept=0)

```
No, those look roughly the same.

Now, let's crunch some more parameter estimates, this time
$\gamma_\text{Dmel}$, as well as define some functions for the reductions due
to RHH and BGS.

```{r}

# nu_bp without the 1/G (since we use L)
dm_gamma <- dm_alpha * dm_m / (2*T)
signif(dm_gamma, 2)

# Hudson and Kaplan (1995) equation 8
HK95 <- function(L, U) exp(-U / L)

RHH_S2N <- function(L, N, gamma, J22) {
  S <- gamma / L * J22
  2*N*S
}

RHH_BGS <- function(theta, L, N, B, gamma, J22) {
  S <- gamma / L * J22
  # generally theta is small, so we approx as theta / (1 + theta) ~ theta
  # (Coop and Ralph 2012 do this, e.g. eqns 18, 19). This allows 
  # for me to treat this as a proportion reduction.
  # theta / (theta + 1/B + 2*N*S)
  theta / (1/B + 2*N*S)
}

```

Quickly, what are the relative strengths?

```{r}

x <- 10^da_ml$log10_diversity / 4e-9
bgs <- 1/(HK95(da_ml$map_length, 1.6))
rhh <- RHH_S2N(da_ml$map_length, x, dm_gamma, dm_J22_Elyashiv)
# plot(x, bgs, col='red', pch=19)
# points(x, rhh, col='green', pch=19)
# plot(x, rhh, col='green', pch=19)

plot(log10(x),bgs/(rhh + bgs), xlab='log10 pop size', pch=19)

```

What about the S&C 2016 stuff?

```{r}

fixT <- function(s, Ne, U) (exp(2*s*Ne) - 1) / (2*U*s*Ne)

SC16_R <- function(Ne, U, L, s) {
  T <- fixT(s, Ne, U)
  exp((-2*U / (exp(2*s) + L))*(1 - 1/(U*T))^3)
}

SC16_R(Ne=dm_Ne, U=dm_U, L=dm_L, s=10^{-1.5}) # s from Elyashiv
HK95(U=dm_U, dm_L)

```

Not much of a difference...

## Diversity Changes Due to Linked Selection

First we try to extrapolate what the effect of linked selection would be in the
Corbett-Detig dataset.

```{r}

# load corbett-detig data, which is data with impact_of_sel
dca <- d[!is.na(d$impact_of_sel), ]
# check against real dataset
dc <- read_tsv('../data/corbett_detig_2015_updated.tsv')  %>%
        mutate(data_source = 'Corbett-Detig et al.',
               range_cd = range) %>%
        rename(diversity = obs_pi)
# should be same size if joins were right
stopifnot(sum(dc$kingdom == 'Animalia')  == nrow(dca))

dca_noself <- dca[dca$selfing != 1, ]

```

First, do we see the same diversity / population size relationship as we do
with the Leffler et al diversity data? Yes:

```{r}

ggplot(dca) + 
  geom_smooth(aes(log10_popsize, log10(diversity)), method='lm', se=FALSE) + 
  geom_point(aes(log10_popsize, log10(diversity))) + 
  geom_text_repel(aes(log10_popsize, log10(diversity), label=species))

```

What does impact of selection's relation with our population size look like? We
confirm a finding of Corbett-Detig et al. with our own N estimate, that impact
of selection increases with N:

```{r}

ggplot(dca_noself) + geom_point(aes(log10_popsize, impact_of_sel)) + 
  geom_smooth(aes(log10_popsize, impact_of_sel), method='lm', se=FALSE) + 
  geom_text_repel(aes(log10_popsize, impact_of_sel, label=species))  + 
  xlab("log10 population size") + ylab("impact_of_sel")

```

What's this linear relationship?

```{r}

save(dca, dca_noself, file="../data/corbett_detig_data.Rdata")

plot(dca$log10_popsize, log10(dca$diversity), pch=19)
cd_div_fit <- lm(log10(diversity) ~ log10_popsize, dca_noself)
abline(cd_div_fit)

plot(dca_noself$log10_popsize, dca_noself$impact_of_sel, 
     pch=19)
ios_fit <- lm(impact_of_sel ~ log10_popsize, dca_noself)
abline(ios_fit)
summary(ios_fit)

```

Fit a simple linear model of the impact of selection / popsize relationship.

```{r}
a <- coef(ios_fit)[1]
b <- coef(ios_fit)[2]

```

This is on a subset of the data with population size, diversity,
and map lengths -- about 40 taxa.

```{r}

# here, we use the complete dataset, which we copy here
dml <- da_cc

dml$ios_ls <- predict(ios_fit, 
                         newdata=tibble(log10_popsize=dml$log10_popsize))
dml$Ne_N_CD15 <- 1-dml$ios_ls

# this uses Ne implied by diversity and μ = 1e-9
dml <- dml %>% 
     # estimate from Elyashiv et al.
     mutate(Ne_N_HK95 = HK95(map_length, U=dm_U)) %>% 
     # mutate(J22 = J22s(10^dm_seldist$log10_s,  dm_seldist$g, dm_Ne, Inf)) %>%
     mutate(J22 = dm_J22_Elyashiv2) %>%

     # both BGS and RHH
     mutate(Ne_N_RHH_BGS = RHH_BGS(1, L=map_length, N=diversity/(4e-9), 
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=J22),
            Ne_N_RHH_BGS_strongsel = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                              B=Ne_N_HK95, gamma=dm_gamma, 
                                              J22=J22 * 10)) %>%
     mutate(Ne_N_HK95_CD15 = Ne_N_HK95 * Ne_N_CD15) %>% 
     mutate(log10_CD15_diversity = log10(diversity / Ne_N_CD15),
            log10_HK95_diversity = log10(diversity / Ne_N_HK95),
            log10_RHH_BGS_diversity = log10(diversity / Ne_N_RHH_BGS),
            log10_RHH_BGS_strongsel_diversity = 
              log10(diversity / Ne_N_RHH_BGS_strongsel),
            log10_diversity = log10(diversity))

# this is the one used in figures -- assumes Ne = Nc
# this got a bit unwieldy as I tried different things... oh well.
dml_full <- da_ml %>%
     # estimate from Elyashiv et al.
     # mutate(J22 = 100*J22s(dm_seldist$s,  dm_seldist$g, 10^log10_popsize, Inf)) %>%
     mutate(J22 = dm_J22_Elyashiv2) %>%
     mutate(Ne_N_HK95 = HK95(map_length, U=dm_U)) %>% 
     mutate(Ne_N_HK95_strongBGS = HK95(map_length, U=10*dm_U)) %>% 
     # both BGS and RHH
     mutate(Ne_N_RHH_BGS = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=J22)) %>%
     mutate(Ne_N_RHH_BGS_strongsel = RHH_BGS(1, L=map_length, N=10^log10_popsize,
                                             B=Ne_N_HK95, gamma=dm_gamma, 
                                             J22=J22 * 10)) %>%

     mutate(Ne_N_RHH_BGS_strongsel_strongBGS = RHH_BGS(1, L=map_length, 
                                                       N=10^log10_popsize,
                                             B=Ne_N_HK95_strongBGS, gamma=dm_gamma, 
                                             J22=J22 * 10)) %>%
     # 10-folder greater Nc
     mutate(Ne_N_RHH_BGS2 = RHH_BGS(1, L=map_length, N=10*10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=J22)) %>%

     # 100-folder greater Nc
     mutate(Ne_N_RHH_BGS3 = RHH_BGS(1, L=map_length, N=100*10^log10_popsize,
                                   B=Ne_N_HK95, gamma=dm_gamma, 
                                   J22=dm_J22_Elyashiv))

save(dml, dml_full, file='../data/linked_sel_div_changes.Rdata')

```

```

What do the reductions using $N_e = 10^6$ (not $N_c$!) for *Drosophila
melanogaster* look like under these parameters for *Drosophila melanogaster*,
under RHH only (no BGS)?

```{r}
# rough approximation, with Ne = 1e6
1-RHH_BGS(1, dm_L, dm_Ne, B=1, gamma=dm_gamma, J22=dm_J22_Elyashiv2)
```

These strong selection reduction, roughly cutting diversity by 66%, is roughly
consistent with Wiehe and Stephan and Elyashiv et al.'s classic sweep only
model (see Figure 6B, CS column).

Using census size, what are the percent reductions in *D. melanogaster*?

```{r}

dml_full %>% filter(species == "Drosophila melanogaster") %>% 
  select(Ne_N_HK95, Ne_N_RHH_BGS)

```

This is a monumental decrease in diversity... it implies we'd be severely
underestimating $\pi_0$.

Ok, now what about the full RHH + BGS model, with *D. melanogaster* $N_e =
10^6$?

```{r}

1-RHH_BGS(1, dm_L, dm_Ne, B=HK95(U=dm_U, L=dm_L), gamma=dm_gamma, J22=dm_J22_Elyashiv2)

```

84% is surprisingly close to Elyashiv et al.'s Figure 6 value of 77% under
BGS+CS, give this simple model.

A draft plot of the effects, red is RHH, blue is BGS. Ultimately I use a
different figure, but this one is still interesting.

```{r}

plot(dml$log10_popsize, dml$log10_diversity, pch=19, 
                  xlim=c(5, 18),
                  ylim=c(-3, 1))

with(dml, segments(log10_popsize, log10_diversity, 
                      log10_popsize, log10_RHH_BGS_diversity, 
                      lty=3, col='red'))
with(dml, segments(log10_popsize, log10_diversity, 
                      log10_popsize, log10_HK95_diversity, 
                      lty=1, col='blue'))

abline(a = log10(4) + -8, b=1)
abline(a = log10(4) + -10, b=1)

```

## IUCN Red List Analysis

Here's a basic analysis of redlist categories and diversity.

```{r}

drl <- da_dps
drl$redlist_cat[is.na(drl$redlist_cat)] <- "NA/DD"
# ordered in terms of endangeredness
iucn_levels <- c("NA/DD", "LC", "NT", "VU", "EN", "CR")
drl$redlist_cat[drl$redlist_cat == 'DD'] <- "NA/DD"
drl$redlist_cat <- factor(drl$redlist_cat, levels=iucn_levels)

drl %>% group_by(redlist_cat) %>% summarize(ntaxa=n())

# use all data, but notice some categories have few species
drl %>% group_by(redlist_cat) %>% summarize(ntaxa=n())

f0_div <- lm(log10_diversity ~ log10_popsize, drl)
f1_div <- lm(log10_diversity ~ redlist_cat + log10_popsize, drl)
f2_div <- lm(log10_diversity ~ redlist_cat, drl)
AIC(f0_div, f1_div, f2_div)
summary(f1_div)
summary(f2_div)

f0_ps <- lm(log10_popsize ~ redlist_cat, drl)
summary(f0_ps)

ggplot(drl, aes(redlist_cat, log10_diversity)) + geom_boxplot()

```

Table for paper:

```{r}

cis <- signif(cbind(coef(f1_div), confint(f1_div)), 2)
xtable(cis)

```

## Session Info

```{r}
sessionInfo()
```