### 2021-02-08 notebooks/stapley_et_al_2017_pcm.Rmd

# Main Analysis of Stapley et al. (2017) Data

```{r, include=FALSE}
options(warnPartialMatchArgs = FALSE,
        warnPartialMatchDollar = FALSE, 
        warnPartialMatchAttr = FALSE)

```

```{r, message=FALSE}

library(ape)
library(phytools)
library(rotl)
library(scales)
library(tidyverse)
library(rstan)
library(loo)
library(coda)
source('../R/utilities.r')

```

## Data Loading

We load the tree from Open Tree of Life downloaded via `rotl` for the taxa in
the Stapley et al. (2017) data. This is for animals, and we've removed all
missing `log10_size` and `map_length` data.

```{r}
load('../data/stapley_et_al_2017_otl.Rdata')
stopifnot(exists('dsatr') && exists('tr'))
```

We lost a few species in matching taxa to the R Open Tree of Life tree; we
quantify that here:

```{r}
total <- suppressMessages(read_tsv('../data/stapley_et_al_2017.tsv')) %>% 
  filter(kingdom %in% 'Animalia') %>%
  nrow()

msg <- "Size of animalia in Stapley et al (2017), total: %d
                            after matching tree: %d"
message(sprintf(msg, total, nrow(dsatr)))
```

## Phylogenetic Tree

Unfortunately, the Open Tree of Life data lacks branch lengths. Like the
original Stapley et al. (2017) analysis, we use unit branch lengths.

```{r, fig.width=6, fig.height=8}
# now, we force an ultrametric tree, just to get the number 
# of edges we need to set to one
utr <- force.ultrametric(tr)

# set all edges to one, as Stapley et al did.
tr$edge.length  <- rep(1, length(utr$edge.length))

# resolve polytomies
trb <- multi2di(tr)
```

What does the tree look like?

```{r}
stopifnot(dsatr$species == tr$tip.label)
nphyla <- length(unique(dsatr$phylum))
phyla_cols <- hue_pal()(nphyla)

# nfamilies <- length(unique(dsatr$family))
# families_cols <- hue_pal()(nfamilies)

# plot(force.ultrametric(trb), cex=0.6, tip.color=phyla_cols[as.factor(dsatr$family)])
plot(force.ultrametric(trb), cex=0.6, tip.color=phyla_cols[as.factor(dsatr$phylum)])

```

## Some Map Length, Body Size, and Chromosome Number EDA

```{r}

dsatr  %>% ggplot(aes(hcn, map_length)) + geom_point() + 
  geom_smooth(method='lm') + xlab('haploid chromosome number') + 
  ylab('map length (Morgans)')

dsatr %>% ggplot(aes(hcn, log10_size)) + geom_point() + 
  geom_smooth() + xlab('body size') + 
  ylab('haploid chromosome number')

dsatr  %>% ggplot(aes(log10_size, hcn)) + geom_point() + 
  geom_smooth(method='lm') + ylab('haploid chromosome number') + 
  xlab('log10(size)')

dsatr %>% ggplot(aes(hcn, map_length/hcn)) + 
  geom_point() + 
  geom_smooth(method='lm') + xlab('haploid chromosome number') + 
  ylab('Morgans per chromosome (map length / haploid chromosome number)')

```

What do number of Morgans / chromosome across the data look like?

```{r}

dsatr %>% mutate(morgans_per_chrom = map_length/hcn) %>% 
  pull(morgans_per_chrom) %>% hist(50, main='Morgans / chromosome')

```

## The Relationship Between Map Length and Body Size

The relationship we'd like to assess is whether map length is positive
correlated with body size; here body size is a proxy for population density,
where they have a negative relationship (e.g. Damuth (1981)).

For quantitative genetic models of linked selection to decouple diversity from
population size, we need a relationship like: $\log(N) = \alpha / L$ where $L$
is map length in Morgans, and $\alpha$ is an unknown positive constant. Since
$\log(N) \propto -\log(S)$ where $S$ is body size, we have: $\log(S) = -\alpha
/ L$, or $L = -\alpha / \log(S)$.

This is the form of the relationship in the Stapley et al (2017), with my body
size data:

```{r}

with(dsatr, plot(map_length, log10_size, pch=19))

```

This obviously does **not** correct for phylogenetic dependencies in the data. 
Currently, I do not see a way to do all of the following things.

1. Keep the units untransformed, as the theory requires the total map length be
in Morgans. For the sake of evalulating the theory, it does not make sense to
calculate the reduction of diversity on a phylogentically corrected value; the
map length in Morgans of a taxa is known.

2. Ideally, a proper PCM would model the evolution of the traits, map length
and body size. Ideally, all the Metazoan taxa included here would have
well-calibrated phylogeny with branch lengths, and we could actually compare
different trait evolution models (early burst, OU, etc) and control for
phylogenetic non-independence more carefully. However, the current phylogeny of
these taxa from the Open Tree of Life has unit branch lengths. Estimating a
phylgoeny for these Metazoan taxa would be outside the scope of this paper.

3. Model the non-linear nature of the data. Ideally, we would have a way to
test whether there's not only a relationship, but a specific non-linear
relationship between the traits measured. But given the lack of branch lengths
(and as far as I know, methods for this) this is not feasible.

Thus, we have two statistical problems: (1) does the body size / map length
trend exist when we remove the effects of pseudoreplication? (2) does the
general trend fit the non-linear relationship needed to decouple diversity?

My plan is two approach (1) assess the linear relationship with PICs, and (2)
find the non-linear least squared fit on both family, and phylum-level averaged
data. I do this first with non-linear least squares, (3) use Bayesian models
(mostly to address issues with the confidence intervals that occurred with the
non-linear least squares fits).

First, we calculate the family and phylum-level averages of the data, used in
both least-squares and Bayesian model fits.

```{r}

# Side note about what I do below: I also take the log of the mean of groups.
# This makes little difference in practice, but *is* different in theory
# (Jensen's inequality). log10_size2 below is doing the
# *incorrect* thing of taking the mean of the values, then the log. This is
# incorrect for the following reason:
#
# x <- rlnorm(100000, 5, 1)
# hist(log(x), 100, col = "grey", border = "grey")
# abline(v=log(mean(x)), col='red')
# abline(v=mean(log(x)), col='black')
#
# The real mean is 5.

family_mean <- dsatr %>% group_by(family) %>%
                 summarize(log10_size=mean(log10_size, na.rm=TRUE), 
                           log10_size2=log10(mean(size, na.rm=TRUE)),
                           map_length=mean(map_length, na.rm=TRUE),
                           hcn=mean(hcn, na.rm=TRUE),
                           n=n())

phyla_mean <- dsatr %>% group_by(phylum) %>%
                 summarize(log10_size=mean(log10_size, na.rm=TRUE), 
                           log10_size2=log10(mean(size, na.rm=TRUE)),
                           map_length=mean(map_length, na.rm=TRUE), 
                           hcn=mean(hcn, na.rm=TRUE),
                           n=n())

```

### Non-linear Least Squares

First, we use the `nls()` function to fit a few different models, and then use
AIC for model comparisons.

```{r}

## WARNING: I label my models differently here than I do in the Bayesian 
## modeling section, so be cautious in assuming what's what.

# family average
mf0 <- nls(log10_size ~ a + b*map_length, data = family_mean, 
         start = list(a=1, b=-2))

mf1 <- nls(log10_size ~ a + b*map_length^power, data = family_mean, 
         start = list(a=1, b=-2, power = -2))

mf2 <- nls(log10_size ~  b*map_length^power, data = family_mean, 
         start = list(b=-2, power = -2))

# phylum average
mp0 <- nls(log10_size ~ a + b*map_length, data = phyla_mean,
           start = list(a=1, b=-2))

mp1 <- nls(log10_size ~ a + b*map_length^power, data = phyla_mean, 
           start = list(a=1, b=1, power = 1),
           control=list(maxiter = 500))

mp2 <- nls(log10_size ~ b*map_length^power, data = phyla_mean, 
           start = list(b=1, power = 1),
           control=list(maxiter = 500))

summary(mp2)
summary(mf2)

BIC(mp0, mp1, mp2)
BIC(mf0, mf1, mf2)

```

With both phylum and family averages, the best fitting model according to BIC
is the non-linear model without an intercept term.

```{r}
# TODO
mf0hcn <- nls(log10_size ~ a + b*hcn, data = family_mean, 
         start = list(a=1, b=-2))

mf1hcn <- nls(log10_size ~ a + b*hcn^power, data = family_mean, 
         start = list(a=1, b=1, power = 1))

mf2hcn <- nls(log10_size ~  b*hcn^power, data = family_mean, 
         start = list(b=-2, power = -2))

```

Let's bootstrap these model fits.

```{r}

best_model <- function(x, start=NULL) {
  if (is.null(start)) {
    start <- list(b=-2, power=-1)
  }
  nls(log10_size ~  b*map_length^power, data = x,
         start=start, trace = FALSE)
}

bootstrap_sample <- function(x) {
  n <- nrow(x)
  x[sample(n, n, replace=TRUE), ]
}

best_model_safe <- possibly(best_model, NULL)

# x sequences for prediction
sf <- seq(0, 53, length.out=100)
sp <- seq(0, 53, length.out=100)

bs_p <- tibble(rep=1:10000) %>%
  # phylum averages
  mutate(fit = map(rep, ~ best_model_safe(bootstrap_sample(phyla_mean)))) %>%
  mutate(keep = !map_lgl(fit, is.null)) %>% filter(keep)

predict_df <- function(fit, x) {
  tibble(x=x, y=predict(fit, newdata=tibble(map_length=x)))
}

bs_p_ci <- bs_p[sample(nrow(bs_p), 5000), ] %>% 
            mutate(pred = map(fit, predict_df, x=sp)) %>% 
            unnest(pred) %>%
            group_by(x) %>% 
            summarize(lower = quantile(y, probs=0.05/2),
                      upper = quantile(y, probs=1-0.05/2))

bs_f <- tibble(rep=1:10000) %>%
  # family averages
  mutate(fit = map(rep, ~ best_model_safe(bootstrap_sample(family_mean)))) %>%
  mutate(keep = !map_lgl(fit, is.null)) %>% filter(keep)

bs_f_ci <- bs_f[sample(nrow(bs_f), 5000), ] %>% 
            mutate(pred = map(fit, predict_df, x=sf)) %>% 
            unnest(pred) %>%
            group_by(x) %>% 
            summarize(lower = quantile(y, probs=0.05/2),
                      upper = quantile(y, probs=1-0.05/2))

# save the averaged data for plots
save(family_mean, phyla_mean, mp2, mf2, 
     sf, sp,
     bs_f_ci, bs_p_ci, file='../data/body_size_map_length.Rdata')

```

Draft plot (see the `notebooks/figures/` directory for actual figure plots).
Now, I prefer the plot below from the HPD intervals.

```{r}

with(family_mean, plot(map_length, log10_size, pch=19, axes=FALSE, cex=0.7,
                       ylab='', xlab='', xlim=c(0, 50), ylim=c(-4.3, 1.1), col='gray48'))

axis(1, seq(0, 50, 10))
axis(2, seq(-4, 1), seq(-4, 1), labels=latex2exp::TeX(sprintf("$10^{%d}$", seq(-4, 1))))

with(bs_p_ci, polygon(c(x, rev(x)), c(lower, rev(upper)), 
                    col=scales::alpha('cornflowerblue', 0.2), lty=0))

with(bs_f_ci, polygon(c(x, rev(x)), c(lower, rev(upper)), 
                    col=scales::alpha('gray42', 0.2), lty=0))

with(phyla_mean, points(map_length, log10_size, pch=19, col='blue', cex=0.7))

#bs_f %>% mutate(pred = map(fit, predict_df, x=sf)) %>%
#  mutate(x=map(pred, ~ lines(.$x, .$y)))
#
#bs_p %>% mutate(pred = map(fit, predict_df, x=sf)) %>%
#  mutate(x=map(pred, ~ lines(.$x, .$y)))

lines(sf, predict(mf2, newdata=tibble(map_length=sf)), lty='dashed', col='gray32')
lines(sp, predict(mp2, newdata=tibble(map_length=sp)), lty='dashed', col='cornflowerblue')

pivot_ci <- function(x, theta, alpha=0.05) 2*theta - rev(quantile(x, probs=c(alpha/2, 1 - alpha/2)))

bs_f %>% mutate(power = map_dbl(fit, ~ coef(.)[2])) %>% 
  pull(power) %>% pivot_ci(coef(mf2)[2])

```

### Bayesian Models -- Family-level Averages

First model is just a simple regression:

$$ log10(\text{size}) ~ \text{Normal}(\alpha + \beta L, \sigma) $$

```{r}

# Stan models use this range for mu_rep and y_rep, used to calculate
# the HPDI and posterior predictive distribution
x_new <- seq(0.1, 53, length.out=100)

bmf0 <- stan(file='../stan/maplength_log10size_linear_family.stan',
             data=list(N = nrow(family_mean), 
                      map_length=family_mean$map_length,
                      log10_size=family_mean$log10_size,
                      x_new = x_new),
             seed=50127839,
             iter=5000, cores=4, control=list(adapt_delta=0.95))

```

Let's assess model convergence:

```{r}

bmf0_pars <- c('alpha', 'beta', 'sigma', 'lp__')
print(bmf0, pars=bmf0_pars)
pairs(bmf0, pars=bmf0_pars)

```

The Rhats show great mixing among chains, we have a high number of effective
samples, and there are no divergences.

Now we look at the next model:

$$ log10(\text{size}) ~ \text{Normal}(\beta L^\gamma, \sigma) $$

(log10 here is a transformation; this is not log normal).

```{r}

bmf1 <- stan(file='../stan/maplength_log10size_power_no_intercept_family.stan',
            data=list(N = nrow(family_mean), 
                      map_length=family_mean$map_length,
                      log10_size=family_mean$log10_size,
                      x_new=x_new),
             seed=59912951,
             iter=10000, cores=4, control=list(adapt_delta=0.95))

bmf1_pars <- c('beta', 'power', 'sigma', 'lp__')
print(bmf1, pars=bmf1_pars)

```

Again, good mixing; let's look at the pairs plot:

```{r}
pairs(bmf1, pars=bmf1_pars)
```

Next, the most complicated model:

$$ log10(\test{size}) ~ \text{Normal}(\alpha + \beta L^\gamma, \sigma) $$

```{r}

bmf2 <- stan(file='../stan/maplength_log10size_power_family.stan',
            data=list(N = nrow(family_mean), 
                      map_length=family_mean$map_length,
                      log10_size=family_mean$log10_size,
                      x_new = x_new),
             seed=125192,
             iter=5000, cores=4, control=list(adapt_delta=0.95))

```

Now, we check the convergence.

```{r}

bmf2_pars <- c('alpha', 'beta', 'power', 'sigma', 'lp__')
print(bmf2, pars=bmf2_pars)
pairs(bmf2, pars=bmf2_pars)

```

Finally, we estimate the coefficients for a model where we decouple N and L
directly, by setting `power = -1`.

```{r}

bmf3 <- stan(file='../stan/maplength_log10size_inverse_family.stan',
            data=list(N = nrow(family_mean), 
                      map_length=family_mean$map_length,
                      log10_size=family_mean$log10_size,
                      x_new = x_new),
             seed=7814,
             iter=5000, cores=4, control=list(adapt_delta=0.95))

bmf3_pars <- c('alpha', 'beta', 'sigma', 'lp__')
print(bmf3, pars=bmf3_pars)

```

Now, we need to compare these models.

```{r}

loow <- function(fit) {
  # a loo() wrapper
  log_lik <- extract_log_lik(fit, merge_chains = FALSE)
  r_eff <- relative_eff(exp(log_lik), cores = 2) 
  loo(log_lik, r_eff = r_eff, cores = 2)
}

loow(bmf0)
loow(bmf1)
loow(bmf2)

loo_compare(loow(bmf0), loow(bmf1), loow(bmf2), loow(bmf3))

```

This shows model 2 (mind they index by 1, not zero) does best, but it's not
significantly different from model 1 (see this answer by the `loo` package's
creator arguing at least a few SDs are necessary to say one model is better
than another).

Looking at WAIC:

```{r}

loo_compare(waic(extract_log_lik(bmf0)),
            waic(extract_log_lik(bmf1)),
            waic(extract_log_lik(bmf2)),
            waic(extract_log_lik(bmf3)))

```

So, again model 2 looks to be the best, though the standard error of the
difference is not very large. Still, arguably we should choose the simpler
model that does best, model 1. This indicates we are not over fitting the power
term compared the linear-terms-only model, which is what we want to see.

### Bayesian Models -- Phylum Averages

Next, we fit these same models on the phylum-level averaged data. Note that
it's worth looking at the Stan files in detail. I had to tweak priors quite a
bit, as an early bad prior choice was leading to divergent transitions, and in
some cases, high `Rhat` values. I still use weakly informative priors, option
for pretty diffuse Student's t distribution for the $\alpha$ and $\beta$
parameters, while using a standard Normal for the power, and a tighter prior on
the standard deviation. This last point deserves some justification. These
phyla-level averages are averages of *many* taxa, and thus, the our prior
expectation of what the standard deviation should be decreases. Early on in
testing, I was using the same prior as the family-level averages. The fits were
very different from the non-linear least squares approach, which was
suspicious. Essentially, we were underfitting the data; our expectation should
be averages of more points have less variance. Changing the priors accordingly
lead to fits more consistent (but still different) from those found with
non-linear least squares. See the Stan model files for more details.

```{r}

bmp0 <- stan(file='../stan/maplength_log10size_linear_phylum.stan',
             data=list(N = nrow(phyla_mean), 
                      map_length=phyla_mean$map_length,
                      log10_size=phyla_mean$log10_size,
                      x_new=x_new),
             seed = 84914,
             iter=10000, cores=4, control=list(adapt_delta=0.99))

bmp1 <- stan(file='../stan/maplength_log10size_power_no_intercept_phylum.stan',
             data=list(N = nrow(phyla_mean), 
                      map_length=phyla_mean$map_length,
                      log10_size=phyla_mean$log10_size,
                      x_new = x_new),
             # increasing adapt_delta decreases the number of divergent transitions
             seed = 41123,
             iter=10000, cores=4, control=list(adapt_delta=0.999))

bmp2 <- stan(file='../stan/maplength_log10size_power_phylum.stan',
                  data=list(N = nrow(phyla_mean), 
                      map_length=phyla_mean$map_length,
                      log10_size=phyla_mean$log10_size,
                      x_new=x_new),
             seed = 874891,
             # we increase the number of iterations and warmup to reduce divergent 
             # transitions
             iter=12000, cores=4, 
             warmup = 5000,
             # increasing adapt_delta decreases the number of divergent transitions
             control=list(adapt_delta=0.999))

bmp3 <- stan(file='../stan/maplength_log10size_inverse_phylum.stan',
            data=list(N = nrow(phyla_mean), 
                      map_length=phyla_mean$map_length,
                      log10_size=phyla_mean$log10_size,
                      x_new = x_new),
             seed=519459,
             iter=5000, cores=4, control=list(adapt_delta=0.95))

```

Next, we let's look at the model output, checking `Rhat` and the number of
effective samples:

```{r}

print(bmp0, pars=bmf0_pars)
print(bmp1, pars=bmf1_pars)
print(bmp2, pars=bmf2_pars)
print(bmp3, pars=bmf3_pars)

```

Next, we look at the pairs plots:

```{r}

pairs(bmp0, pars=bmf0_pars)
pairs(bmp1, pars=bmf1_pars)
pairs(bmp2, pars=bmf2_pars)
pairs(bmp3, pars=bmf2_pars)

```

Again, all of these look reasonable. Notice there are no (or very, very few
divergent transitions). All `Rhat` values are < 1.

Next, let's look at the estimated expected log posteriors:

```{r}
print(loow(bmp0))
print(loow(bmp1))
print(loow(bmp2))
print(loow(bmp3))
```

And we compare them, using `loo`'s compare function and WAIC:

```{r}

loo_compare(loow(bmp0), loow(bmp1), loow(bmp2), loow(bmp3))

loo_compare(waic(extract_log_lik(bmp0)),
            waic(extract_log_lik(bmp1)),
            waic(extract_log_lik(bmp2)),
            waic(extract_log_lik(bmp3)))

```

These indicate that in terms of expected log predictive density, models 1
(again, mind the fact my model numbers are zero-indexed and theirs are
1-indexed), is much better than the linear model.

Next, let's look at the HPD intervals:

```{r}

maplength_bodysize_plot(phyla_mean)
ci_polygon(x_new, bmp1, color='gray', type='hpdi', par='mu_rep')
ci_polygon(x_new, bmp2, color='orange', type='hpdi', par='mu_rep')
ci_polygon(x_new, bmp3, color='red', type='hpdi', par='mu_rep')

```

For models 1 and 2, what do the posterior predictive distributions look like?

```{r}

maplength_bodysize_plot(family_mean)
ci_polygon(x_new, bmf2, color='gray', type='quantile', par='y_rep')
ci_polygon(x_new, bmf1, color='green', type='quantile', par='y_rep')

```

So, we output these models

```{r}

save(family_mean, phyla_mean, 
     bmf1, bmp1,
     bmf2, bmp2,
     bmf3, bmp3,
     x_new,
     file='../data/body_size_map_length_bayesian.Rdata')

```

## Phylogenetic Independent Contrast

Next, I want to establish that the relationship between map length and body
size is significant with phylogenetic corrections. Again, these results would
be greatly improved by having non-unit branch lengths. Given that unit branch
lengths probably over-corrects for phylogenetic signal over these long
evolutionary timeframes, this is a conservative approach.

```{r}
# TODO
plot(trb)

plotTree(force.ultrametric(tr))
# remove polytomies
utrb <- multi2di(tr)

log_ml <- log10(dsatr$map_length)
log_size <- dsatr$log10_size

log_ml_pic = as_tibble(pic(log_ml, utrb, var.contrasts=TRUE))
log_size_pic = as_tibble(pic(log_size, utrb, var.contrasts=TRUE))

plot(sqrt(log_ml_pic$variance), abs(log_ml_pic$contrasts))
plot(sqrt(log_size_pic$variance), abs(log_size_pic$contrasts))

m_pic <- lm(log_size_pic$contrasts ~ log_ml_pic$contrasts)
summary(m_pic)

plot(log_ml_pic$contrasts, log_size_pic$contrasts)
abline(m_pic)

plot(a$contrasts, b$contrasts)
abline(lm(a$contrasts ~ b$contrasts))
m <- lm(a$contrasts ~ b$contrasts)

dsic <- tibble(maplength_contrast=a$contrasts, size_contrast=b$contrasts)

fit_contrast_lm <- function(d) {
  n <- nrow(d)
  lm(size_contrast ~ maplength_contrast, data=d[sample(n, n, replace=TRUE), ])
}

bs <- tibble(rep = 1:5000) %>% 
  mutate(fit = map(rep, ~ fit_contrast_lm(dsic))) %>% 
  mutate(pval = map_dbl(fit, ~ coef(.)[2]))

alpha <- 0.05
quantile(bs$pval, c(alpha/2, 1-alpha/2))

bs %>% ggplot(aes(x=pval)) + geom_histogram()

par(opar)
```

## Diversity and Map Length

First, we merge in all the animal data we have.

```{r}

# load in the Corbett-Detig et al (2015) data
dc <- read_tsv('../data/corbett_detig_2015_updated.tsv') %>% 
         filter(kingdom == 'animal') %>%
         select(species, size, diversity=obs_pi, map_length)

lds <- read_tsv('../data/leffler_et_al_2012_updated.tsv') %>%
         filter(kingdom == 'Animalia') %>%
         select(species, size, diversity)

# The Stapley data has size (that I've added) and map length, but not diversity.
# The Corbett-Detig data has all three. 
# The Leffer data has size (that I've added) and diversity, but not map length.
dsatr2 <- dsatr %>% select(species, size, map_length)

dt <- lds %>% left_join(dsatr2)  %>% 
        filter(!is.na(size), !is.na(diversity), !is.na(map_length)) %>%
        # we join in the Corbett-Detig data
        full_join(dc, by=c('species'))

```

This last join didn't merge like columns. Which is good, because they allow us
to do some quality control (in the good sense, not the Iowa Caucus sense).

```{r}

ggplot(dt, aes(diversity.x, diversity.y)) + geom_point() +
  geom_abline(yintercept=0, slope=1, color='gray')

ggplot(dt, aes(size.x, size.y)) + geom_point() + ylim(0, 0.8) +
  geom_abline(yintercept=0, slope=1, color='gray')

ggplot(dt, aes(map_length.x, map_length.y/100)) + geom_point() + 
  geom_abline(yintercept=0, slope=1, color='gray')

```

The values all look good. I'll average them for the final dataframe:

```{r}

merge_average <- function(x, y) rowMeans(cbind(x, y), na.rm=TRUE)

# final dataframe
dtf <- dt %>% mutate(size = map2_dbl(size.x, size.y, merge_average),
              diversity = map2_dbl(diversity.x, diversity.y, merge_average),
              # the map_length.y / 100 below is converting cM in M
              map_length = map2_dbl(map_length.x, map_length.y/100, merge_average)) %>%
  select(-size.x, -size.y, -diversity.x, -diversity.y, -map_length.x, -map_length.y)

```

Now, we use the SC98 and HK94 reductions in $N_e$ to calculate the shift. This
data creates Figure 2. For SC98, we use $\alpha = 1$, where $\alpha = C^2 /
(1-Z)$. This value is greater than we'd expect under Drosophila. For HK94, we
use the Elyashiv et al data.

$$
\pi = 4 N \mu \exp\{- U / L\}
$$

$$
\pi = \pi_0 \exp\{- U / L\}
$$

Elyashiv et al report the average reduction, which is the ratio of $\bar{\pi} /
\pi_0$. They report under the BG+CS model an average reduction of 77% to 89%.

$$
\pi = \pi_0 \exp\{- U / L\}
$$

```{r}

SC98 <- function(L, alpha) exp(-alpha /L)
HK94 <- function(L, U) exp(-U / L)

dtft <- dtf %>%  
  filter(species != 'Ciona savignyi') %>%
  mutate(Ne_N_SC98 = SC98(map_length, alpha=1))  %>%
  mutate(Ne_N_HK94 = HK94(map_length, U=1.6)) # estimate from Elyashiv et al.

dtft %>%
  filter(species != 'Ciona savignyi') %>%
  mutate(decoupled_diversity = size) %>%
  ggplot(aes(-log10(size), (diversity))) + geom_point()  + 
   # geom_smooth(aes(-log10(size), decoupled_diversity), color='red', method='lm') + 
  geom_segment(aes(-log10(size), (diversity), 
                   xend=-log10(size), yend=(diversity / Ne_N_HK94))) + 
  geom_segment(aes(-log10(size), (diversity), 
                   xend=-log10(size), yend=(diversity / Ne_N_SC98)), color='red') +
  scale_y_log10()

save(dtft, file='../data/diversity_size_theory.Rdata')
```

## Fecundity and body size

Next, I look at the relationship between fecundity and body size. If we imagine
a Poisson number of offspring, the mean level of fecundity is equal to the
variance; this is a conservative model, as overdispersion is very likely.

Note that we violate a key assumption from the start: independence between
points. Each of these points are taxa, so they are phylogenetically
non-independent (Felsenstein, 1985). Knowing this is a serious problem, I still
first fit this relationship as a simple linear model, which illustrates some
pathologies with model fit.

```{r}

dr <- read_tsv('../data/romiguier_et_al_2014_updated.tsv') %>%
        mutate(log10_fecundity = log10(fecundity)) %>%
        #filter(fecundity < 1e5) %>% # optional remove an outlier
        filter(!is.na(log10_fecundity), !is.na(log10_size))

size_fecundity_plot <- function() {
  with(dr %>% filter(fecundity < 8e4),
       plot(log10_size, log10_fecundity, 
            axes=FALSE, 
            pch=19,
            xlim=c(-3.1, 1),
            col='gray42'))
  axis(1, seq(-3, 1), labels=latex2exp::TeX(sprintf("$10^{%d}$", seq(-3, 1))))
  axis(2, seq(-3, 5), labels=latex2exp::TeX(sprintf("$10^{%d}$", seq(-3, 5))))
}

size_fecundity_plot()

lmf <- lm(log10_fecundity ~ log10_size, dr)
abline(lmf)

```

We see two things: (1) a great deal of heteroscedasticity, (2) a poor fit
around the right end of the plot (all points are below the line of best fit).
Looking at a plot of the residuals vs fitted:

```{r}
plot(lmf, which=1)
```

This pathology is also reflected in the QQ-plot:

```{r}
plot(lmf, which=2)
```

So instead, I will try to fit this with a Bayesian model using Stan. As before,
I ameliorate non-independence by considering family and phylum-level averages.

```{r}

dr_family_averages  <- dr %>% group_by(family) %>% 
  summarize(log10_size = mean(log10_size, na.rm=TRUE),
            log10_fecundity=mean(log10_fecundity, na.rm=TRUE),
            fecundity=mean(fecundity, na.rm=TRUE),
            n=n())

dr_phylum_averages  <- dr %>% group_by(phylum) %>% 
  summarize(log10_size = mean(log10_size, na.rm=TRUE),
            log10_fecundity=mean(log10_fecundity, na.rm=TRUE),
            fecundity=mean(fecundity, na.rm=TRUE),
            n=n())

```

We first fit a log-normal regression model on the family-level averaged data.

```{r}

lnf_xnew <- seq(-3, 1, length.out=100)

lnf <- stan(file='../stan/fecundity_log10size.stan',
            data=list(N = nrow(dr_family_averages), 
                      log10_fecundity=dr_family_averages$log10_fecundity,
                      log10_size=dr_family_averages$log10_size,
                      x_new = lnf_xnew),
            seed = 861234,
            iter=5000, cores=4, control=list(adapt_delta=0.95))

```

Looking at model diagnostics, we see good number of effective samples, and
`Rhat`, showing good mixing of the chains.

```{r}
lnf_pars <- c('beta', 'sigma', 'lp__')
print(lnf, pars=lnf_pars)
```

And finally, in our pair plot, we see no divergent transitions.

```{r}
pairs(lnf, pars=lnf_pars)
```

Next, I do some posterior predictive checks, visualizing it along side the
HPDI. Overall the model fit is good.

```{r}

size_fecundity_plot()
ci_polygon(lnf_xnew, y=extract(lnf, pars='mu_rep')$mu_rep,
           color='gray', type='hpdi')
ci_polygon(lnf_xnew, y=extract(lnf, pars='y_rep')$y_rep,
           color='red', type='quantile', transp=0.1)
lnf_p <- colMeans(as_tibble(extract(lnf, pars='mu_rep')))
lines(lnf_xnew, lnf_p, col=alpha('black', 0.4), lwd=3)

```

We save these data for prettier publication-quality plots.

```{r}
save(dr,
     dr_family_averages, dr_phylum_averages,
     lnf_xnew, lnf, 
     file='../data/body_size_fecundity.Rdata')
```

## Fecundity and Body Size

Now, we look at the relationship between fecundity, body size, and diversity.

```{r}

dr %>% ggplot() + geom_point(aes(log10_size, log10(piS))) + 
   geom_segment(aes(log10_size, log10(piS), 
                    xend=log10_size, yend=log10(piS*(fecundity))))

```